A survey of registration practices among observational researchers using preexisting datasets

### 2023-04-04 manuscript/manuscript.Rmd
Code check (Robby)
removed participants, edited codebook generation, minor tweaks to comments.

```{r loadingData}
processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
```

@@

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).

@@

# Calculate the fequency and proportion of responses for each response level
confirmatory <- level_percentage(
  processed,
  method_confirmatory,
  method_levels,
  exclude_missing = FALSE
) %>%
  filter(method_confirmatory %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
  summarise(
    n = sum(n),
    # Since n_sum should be the same for the whole dataset we can use unique to return one value
    n_sum = unique(n_sum),
    percentage = sum(percentage))

>>>>

```{r loadingData}
processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
processed_all <- read_csv(here::here("data/processed/ecaw_processed_all_data.csv"))
```

@@

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed) ` times, leading to a response rate of `r round(nrow(processed_all) / 1094 * 100)`% for complete surveys and `r round((nrow(processed_all) - nrow(processed)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).

@@

# Calculate the frequency and proportion of responses for each response level
confirmatory <- level_percentage(
  processed,
  method_confirmatory,
  method_levels,
  exclude_missing = FALSE
) %>%
  filter(method_confirmatory %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
  summarise(
    n = sum(n),
    # Since n_sum should be the same for the whole dataset we can use unique to return one value
    n_sum = unique(n_sum),
    percentage = sum(percentage))

### 2023-04-04 manuscript/manuscript.tex
Code check (Robby)
removed participants, edited codebook generation, minor tweaks to comments.

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn't been active in research for 30 years. Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.22 minutes (IQR: 4.30 to 13.10).

Respondents published a median of NA (IQR 2 to 25) studies using preexisting observational data (Figure S1). They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}. 61\% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% reported being less concerned.

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.''} (72\%; 77/107) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 84/107) (Figure 2, top panel). At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 2, bottom panel).

@@

\caption{\textbf{Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. For the bottom panel we excluded the missing responses (n = 2; 2) and responses of ``I don't understand the question'' (n = 4; 3).}\label{fig:typicallyEcawPlot}
\end{figure}

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2A). About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107) (Figure 3). 75\% (80/107) reported that they never or almost never blind the data analyst (Figure 3). Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes (Figure 3).

@@

25\% (26/102) of respondents agreed (versus 44\%; 45/102 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). 53\% (50/95) agreed (20\%; 19/95 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 55\% (53/97) agreed (10\%; 10/97 disagreed) that ALSPAC should run a study on ECAWs. 45\% (43/95) agreed (22\%; 21/95 disagreed) that they would prefer using an ECAW than using typical preregistration.

@@

\caption{\textbf{Responses to survey questions about using ECAWs.} The 4 questions we excluded missing values (\emph{n} = 3; 3; 3; 3), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
\end{figure}

@@

\hypertarget{supplementary-materials-c}{%
\section{Supplementary materials C}\label{supplementary-materials-c}}

@@

\caption{\label{tab:languageTable}What programming language or software do you use for your analyses of preexisting observational data?}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrr}
\toprule
\textbf{Programming language} & \textbf{N} & \textbf{Percentage of respondents}\\
\midrule
R & 65 & 61\\
\cmidrule{1-3}
Stata & 48 & 45\\
\cmidrule{1-3}
SPSS & 17 & 16\\
\cmidrule{1-3}
SAS & 15 & 14\\
\cmidrule{1-3}
Python & 6 & 6\\
\cmidrule{1-3}
Missing & 4 & 4\\
\cmidrule{1-3}
Mplus & 3 & 3\\
\cmidrule{1-3}
Bash & 2 & 2\\
\cmidrule{1-3}
MATLAB & 1 & 1\\
\cmidrule{1-3}
Nextflow & 1 & 1\\
\cmidrule{1-3}
plink2 & 1 & 1\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed 103 times and partially completed 20 times, leading to a response rate of 11\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn't been active in research for 30 years. Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.40 minutes (IQR: 4.60 to 13.10).

Respondents published a median of NA (IQR 2 to 26.20) studies using preexisting observational data (Figure S1). They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}. 62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% (6/100) reported being less concerned.

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.''} (72\%; 74/103) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 81/103) (Figure 2, top panel). At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 2, bottom panel).

@@

\caption{\textbf{Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. For the bottom panel we excluded the missing responses (n = ; ) and responses of ``I don't understand the question'' (n = 3; 2).}\label{fig:typicallyEcawPlot}
\end{figure}

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2A). About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103) (Figure 3). 77\% (79/103) reported that they never or almost never blind the data analyst (Figure 3). Almost all respondents answered that they use both exploratory (93\%; 96/103) and confirmatory (87\%; 90/103) analyses at least sometimes (Figure 3).

@@

26\% (26/101) of respondents agreed (versus 45\%; 45/101 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). 53\% (50/94) agreed (20\%; 19/94 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 55\% (53/96) agreed (10\%; 10/96 disagreed) that ALSPAC should run a study on ECAWs. 46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using typical preregistration.

@@

\caption{\textbf{Responses to survey questions about using ECAWs.} For the 4 questions we excluded missing values (\emph{n} = 0; 0; 0; 0), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
\end{figure}

@@

@@

\caption{\label{tab:languageTable}What programming language or software do you use for your analyses of preexisting observational data?}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrr}
\toprule
\textbf{Programming language} & \textbf{N} & \textbf{Percentage of respondents}\\
\midrule
R & 65 & 63\\
\cmidrule{1-3}
Stata & 48 & 47\\
\cmidrule{1-3}
SPSS & 17 & 17\\
\cmidrule{1-3}
SAS & 15 & 15\\
\cmidrule{1-3}
Python & 6 & 6\\
\cmidrule{1-3}
Mplus & 3 & 3\\
\cmidrule{1-3}
Bash & 2 & 2\\
\cmidrule{1-3}
MATLAB & 1 & 1\\
\cmidrule{1-3}
Nextflow & 1 & 1\\
\cmidrule{1-3}
plink2 & 1 & 1\\
\bottomrule
\end{tabular}}
\end{table}
